Robust Estimation, Prediction and Control with \\ Continuous States and Dynamics Structure

### 2020-01-14 main.tex
Remove regularization matrix \Sigma_d

\paragraph{Regularised least square} Since we expect that $\Phi_n$ may have zero-coordinates during possibly long time intervals (i.e. some features are not always active), we consider the weighted $L_2$-regularised regression problem with weights  $\Sigma_p\in\Real^{p\times p}$ and $\Sigma_d\in\Real^{d\times d}$ and parameter $\lambda\in\Real^+_*$:

\begin{equation}
    \label{eq:regression_min}
    \min_{\theta\in\Real^d} \sum_{n=1}^N \|y_n -\Phi_n\theta\|_{\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2
\end{equation}

@@

\begin{theorem}[Regularised solution]
\label{thm:regularized_solution}
The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} y_n
\end{equation}
where 
\begin{equation*}
    G_{N, \lambda} = \sum_{n=1}^N \Phi_{n}^\transp\Sigma_p^{-1}\Phi_{n}  + \lambda \Sigma_d \in \Real^{d\times d}.
\end{equation*}
\end{theorem}

Denote $\eta(t) = C\nu(t) + D\omega(t)$. By substituting the expression of $y_n$ into \eqref{eq:vector_rls}, we obtain:
\begin{lemma}[Regression error]
\label{lem:regression-error}
\begin{align*}
    \theta_{N,\lambda} - \theta = G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n - \lambda G_{N, \lambda}^{-1}\Sigma_d\theta 
\end{align*}
\end{lemma}

@@

Then, 
\begin{align}
\|\theta_{N,\lambda} - \theta\|_1 \leq &\|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty \\
&+ \lambda\|G_{N, \lambda}^{-1}\Sigma_d\|_1\|\theta\|_\infty \nonumber
\end{align}

@@

\begin{proof}
We differentiate $J(\theta) = \sum_{n=1}^N \|y_n -\Phi_n\theta\|_{\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{\Sigma_d}^2$ as in  \eqref{eq:regression_min} with respect to $\theta$:

\begin{align*}
    \nabla_{\theta} J(\theta) &= \sum_{n=1}^N\nabla_{\theta} (y_n - \Phi_n\theta)^\transp\Sigma_p^{-1}(y_n - \Phi_n\theta) + \nabla_{\theta} \lambda\|\theta\|_{\Sigma_d}^2\\
    &= -2\sum_{n=1}^N y_n^\transp\Sigma_p^{-1}\Phi_n + 2\sum_{n=1}^N\theta^\transp(\Phi_n^\transp\Sigma^{-1}\Phi_n) +  2 \lambda \theta^\transp \Sigma_d
\end{align*}

Hence,
\begin{align*}
    \nabla_{\theta} J(\theta) = 0 \iff \left(\sum_{n=1}^N\Phi_n^\transp\Sigma_p^{-1}\Phi_n + \Sigma_d\right)\theta = \sum_{n=1}^N y_n^\transp\Sigma_p^{-1}\Phi_n
\end{align*}
\end{proof}

@@

\begin{proof}
In order to study the concentration of $\theta_{N,\lambda}$, we substitute the expression of $y_n$ into \eqref{eq:vector_rls}:
\begin{align*}
    \theta_{N,\lambda} &= G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} (\Phi_n\theta +\eta_n)\\
    &= \theta - G_{N, \lambda}^{-1}\lambda\Sigma_d\theta + G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n
\end{align*}

@@

% The following uses the method of mixtures.
% Let $\Gamma$ be a Gaussian random variable which is independent of all other random variables and whose covariance is $\Sigma^{-1}$. (in our estimator, $\Sigma = \lambda\Sigma_d$).

@@

Hence, for all $x\in\Real^d$,
\begin{align*}
    x^\transp\theta_{N,\lambda}  -x^\transp\theta &= x^\transp G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n
    - \lambda x^\transp G_{N, \lambda}^{-1}\Sigma_d\theta\\
    &= \inp{x}{\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n}_{G_{N, \lambda}^{-1}} - \lambda\inp{x}{\Sigma_d\theta}_{G_{N, \lambda}^{-1}}
\end{align*}

Using the Cauchy-Schwartz inequality, we get:
\begin{align*}
    |x^\transp\theta_{N,\lambda}  -x^\transp\theta| &\leq \|x\|_{G_{N, \lambda}^{-1}}\left(\left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\Sigma_d\theta\|_{G_{N, \lambda}^{-1}}\right)
\end{align*}

In particular, for $x = G_{N,\lambda}(\theta_{N,\lambda} - \theta)$, we get after simplifying with $\| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}}$:
\begin{align*}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} &\leq \left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\Sigma_d\theta\|_{G_{N, \lambda}^{-1}}
\end{align*}

By applying \autoref{prop:concentration}, we obtain:
\begin{align*}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} &\leq 
     + \lambda\|\Sigma_d\theta\|_{G_{N, \lambda}^{-1}}
\end{align*}
\end{proof}

>>>>

\paragraph{Regularised least square} Since we expect that $\Phi_n$ may have zero-coordinates during possibly long time intervals (i.e. some features are not always active), we consider the weighted $L_2$-regularised regression problem with weights  $\Sigma_p\in\Real^{p\times p}$ and parameter $\lambda\in\Real^+_*$:

\begin{equation}
    \label{eq:regression_min}
    \min_{\theta\in\Real^d} \sum_{n=1}^N \|y_n -\Phi_n\theta\|_{\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{}^2
\end{equation}

@@

\begin{theorem}[Regularised solution]
\label{thm:regularized_solution}
The solution to \eqref{eq:regression_min} is:
\begin{equation}
    \label{eq:vector_rls}
    \theta_{N,\lambda} = G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} y_n
\end{equation}
where 
\begin{equation*}
    G_{N, \lambda} = \sum_{n=1}^N \Phi_{n}^\transp\Sigma_p^{-1}\Phi_{n}  + \lambda I_d \in \Real^{d\times d}.
\end{equation*}
\end{theorem}

Denote $\eta(t) = C\nu(t) + D\omega(t)$. By substituting the expression of $y_n$ into \eqref{eq:vector_rls}, we obtain:
\begin{lemma}[Regression error]
\label{lem:regression-error}
\begin{align*}
    \theta_{N,\lambda} - \theta = G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n - \lambda G_{N, \lambda}^{-1}\theta 
\end{align*}
\end{lemma}

@@

Then, 
\begin{align}
\|\theta_{N,\lambda} - \theta\|_1 \leq &\|G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\|_1\|\eta\|_\infty \\
&+ \lambda\|G_{N, \lambda}^{-1}\|_1\|\theta\|_\infty \nonumber
\end{align}

@@

\begin{proof}
We differentiate $J(\theta) = \sum_{n=1}^N \|y_n -\Phi_n\theta\|_{\Sigma_p^{-1}}^2 + \lambda\|\theta\|_{}^2$ as in  \eqref{eq:regression_min} with respect to $\theta$:

\begin{align*}
    \nabla_{\theta} J(\theta) &= \sum_{n=1}^N\nabla_{\theta} (y_n - \Phi_n\theta)^\transp\Sigma_p^{-1}(y_n - \Phi_n\theta) + \nabla_{\theta} \lambda\|\theta\|_{}^2\\
    &= -2\sum_{n=1}^N y_n^\transp\Sigma_p^{-1}\Phi_n + 2\sum_{n=1}^N\theta^\transp(\Phi_n^\transp\Sigma^{-1}\Phi_n) +  2 \lambda \theta^\transp
\end{align*}

Hence,
\begin{align*}
    \nabla_{\theta} J(\theta) = 0 \iff \left(\sum_{n=1}^N\Phi_n^\transp\Sigma_p^{-1}\Phi_n + I_d\right)\theta = \sum_{n=1}^N y_n^\transp\Sigma_p^{-1}\Phi_n
\end{align*}
\end{proof}

@@

\begin{proof}
In order to study the concentration of $\theta_{N,\lambda}$, we substitute the expression of $y_n$ into \eqref{eq:vector_rls}:
\begin{align*}
    \theta_{N,\lambda} &= G_{N, \lambda}^{-1} \sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1} (\Phi_n\theta +\eta_n)\\
    &= \theta - G_{N, \lambda}^{-1}\lambda\theta + G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n
\end{align*}

@@

% The following uses the method of mixtures.
% Let $\Gamma$ be a Gaussian random variable which is independent of all other random variables and whose covariance is $\Sigma^{-1}$. (in our estimator, $\Sigma = \lambda$).

@@

Hence, for all $x\in\Real^d$,
\begin{align*}
    x^\transp\theta_{N,\lambda}  -x^\transp\theta &= x^\transp G_{N, \lambda}^{-1}\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n
    - \lambda x^\transp G_{N, \lambda}^{-1}\theta\\
    &= \inp{x}{\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n}_{G_{N, \lambda}^{-1}} - \lambda\inp{x}{\theta}_{G_{N, \lambda}^{-1}}
\end{align*}

Using the Cauchy-Schwartz inequality, we get:
\begin{align*}
    |x^\transp\theta_{N,\lambda}  -x^\transp\theta| &\leq \|x\|_{G_{N, \lambda}^{-1}}\left(\left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\theta\|_{G_{N, \lambda}^{-1}}\right)
\end{align*}

In particular, for $x = G_{N,\lambda}(\theta_{N,\lambda} - \theta)$, we get after simplifying with $\| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}}$:
\begin{align*}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} &\leq \left\|\sum_{n=1}^N \Phi_n^\transp \Sigma_p^{-1}\eta_n\right\|_{G_{N, \lambda}^{-1}} + \lambda\|\theta\|_{G_{N, \lambda}^{-1}}
\end{align*}

By applying \autoref{prop:concentration}, we obtain:
\begin{align*}
    \| \theta_{N,\lambda}  - \theta\|_{G_{N,\lambda}} &\leq 
     + \lambda\|\theta\|_{G_{N, \lambda}^{-1}}
\end{align*}
\end{proof}
